Supporting Information: Thermal sensitivity across forest vertical profiles: patterns, mechanisms, and ecological implications

### 2021-09-14 vertical-thermal-review.Rmd
Update vertical-thermal-review.Rmd

**Canopy trees account for the majority of forest ecosystem water and carbon cycling.**
While studies partitioning transpiration across forest vertical gradients are rare, both ET and transpiration have been shown to increase with height in a *Picea abies* forest in Germany, such that the upper half of the canopy contributed an estimated 80% of daytime ET [@staudtVerticalStructureEvapotranspiration2011]. Similarly, in a tropical forest in the Brazilian Amazon, canopy and subcanopy trees jointly contributed approximately 93% of ET, or 88% of transpiration [Fig `r fig_ecosystem`a; @kunertRevisedHydrologicalModel2017], and trees >33 cm DBH contributed >80% of transpiration in a forest in Costa Rica [@mooreUpscalingTranspirationDiverse2018]. In terms of C cycling, it has been estimated that canopy strata contribute $\geq$ 64% of net daytime ~CO~2~ uptake [i.e., GPP - ecosystem respiration, including from soil; @missonPartitioningForestCarbon2007]. Large trees also dominate in terms of woody aboveground net primary productivity ($ANPP_{woody}$) and mortality ($M_{woody}$), with trees $\geq$ 10cm DBH usually contributing > 85% of ANPPwoody and Mwoody across 25 globally distributed forests [Fig. `r fig_ecosystem`b, @piponiot_sizerelated_inreview].

>>>>

**Canopy trees account for the majority of forest ecosystem water and carbon cycling.**
While studies partitioning transpiration across forest vertical gradients are rare, both ET and transpiration have been shown to increase with height in a *Picea abies* forest in Germany, such that the upper half of the canopy contributed an estimated 80% of daytime ET [@staudtVerticalStructureEvapotranspiration2011]. Similarly, in a tropical forest in the Brazilian Amazon, canopy and subcanopy trees jointly contributed approximately 93% of ET, or 88% of transpiration [Fig `r fig_ecosystem`a; @kunertRevisedHydrologicalModel2017], and trees >33 cm DBH contributed >80% of transpiration in a forest in Costa Rica [@mooreUpscalingTranspirationDiverse2018]. In terms of C cycling, it has been estimated that canopy strata contribute $\geq$ 64% of net daytime CO~2~ uptake [i.e., GPP - ecosystem respiration, including from soil; @missonPartitioningForestCarbon2007]. Large trees also dominate in terms of woody above ground net primary productivity ($ANPP_{woody}$) and mortality ($M_{woody}$), with trees $\geq$ 10cm DBH usually contributing > 85% of ANPPwoody and Mwoody across 25 globally distributed forests [Fig. `r fig_ecosystem`b, @piponiot_sizerelated_inreview].
